---
title: "Report"
author: "Mason Wong, Eva Yin, Zhuolin Jiang, Matthew Shu"
date: "14/05/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
```

```{r}
library(GGally)
library(tidyverse)
library(visdat)
library(MASS) # for the forward, backward and bidirectional search
```

## Data Description & Visualisation

To Do:

* Summary Statistics
* Check if some relationships are curvilinear
* Correlation Matrix <- discuss multicollinearity < `qtlcharts::iplotCorr(df)` for innovation, `ggally` for all pairs
* Correlation should also give us hints as to what interaction effects to fit
* Missing Data / Strange Data <- relevant to model building <- vismiss for innovation
* Some background / literature
* Literature guidance on important variables for model from domain knowledge

https://stats.stackexchange.com/questions/395562/interpreting-interaction-term-on-highly-correlated-variables?rq=1

https://www.theanalysisfactor.com/regression-modelshow-do-you-know-you-need-a-polynomial/

```{r}
df = read.csv("../data/seatbelt_group_14.csv", stringsAsFactors = TRUE)
df %>% head(. , 3)
```

 

* It seems like data is in the range of 1983 to 1997
* There is a lot of missing data surrounding the rate of seatbelt usage in each state, in each year.
* There is no missing data elsewhere. 
* Potential solution might be to build a model without the seatbelt & w/ seatbelt to avoid dropping all the rows

```{r}

# Get an initial feel of our data
vis_dat(df)
str(df)

# a temporary data frame, made to observe the NA's specifically across each state at each year
temp = df %>% 
  dplyr::select(state, year, seatbelt) %>% 
  group_by(year) %>% 
  arrange(year) %>% 
  arrange(state)

# for the NA's replace them with zeroes
temp = temp %>% mutate(seatbelt.no.na = case_when(is.na(seatbelt) ~ 0, TRUE ~ seatbelt))

# define function for below 
every_nth = function(n) {
  return(function(x) {x[c(TRUE, rep(FALSE, n - 1))]})
}
# observe the plot of NA's by year and state
temp %>% ggplot(. , aes(x = factor(year), y = seatbelt.no.na)) + 
  geom_point() + 
  theme(axis.text.x = element_text(angle = 60, vjust = 0.5)) +
  scale_x_discrete(breaks = every_nth(n = 3)) + 
  xlab('year') + 
  ylab('self reported seatbelt proportions') + 
  facet_wrap(~state)
```

High collinearity

* `Income` and `Year` (Positive)
* `Income` and `Fatalities` (Negative)
* `Fatalities` and `Year` (Negative)
* `Speed65`, `Speed75` more prevalent in later years

```{r}
df %>%
  dplyr::select(-seatbelt, -state) %>%
  ggpairs()
#ggpairs(df)

```


## Coarse grain search with forward, backward and bidirectional 
We firstly get rid of the variables to do with state and seatbelt
```{r}
# take away state and seatbelt as variables and create the full and null model
df.no_ss = df %>% dplyr::select(-state, -seatbelt)
n = nrow(df.no_ss)
full_model = lm(fatalities ~ . , data = df.no_ss)
null_model = lm(fatalities ~ 1 , data = df.no_ss)
```

we perform a forward search with the aic and bic criterion
```{r}
# forward model with aic criterion
forward_model_aic = stepAIC(null_model, 
                        scope = list(upper = formula(full_model),
                                     lower = formula(null_model)),
                        direction = 'forward',
                        k = 2)
# forward model with bic criterion
forward_model_bic = stepAIC(null_model, 
                        scope = list(upper = formula(full_model),
                                     lower = formula(null_model)),
                        direction = 'forward',
                        k = log(n))
```
$$ fatalities \thicksim income + age + speed70 + alcohol$$

We perform a backward search
```{r}
# backward model with AIC criterion
backward_model_AIC = stepAIC(full_model, 
                        scope = list(upper = formula(full_model),
                                     lower = formula(null_model)),
                        direction = 'backward',
                        k = 2)
# backward model with BIC criterion
backward_model_BIC = stepAIC(full_model, 
                        scope = list(upper = formula(full_model),
                                     lower = formula(null_model)),
                        direction = 'backward',
                        k = log(n))
```
$$fatalities \thicksim speed70 + alcohol + income + age$$

We perform a bidirectional search

```{r}
# bidirectional model with AIC criterion
bidirectional_model_AIC = stepAIC(full_model, 
                        scope = list(upper = formula(full_model),
                                     lower = formula(null_model)),
                        direction = 'both',
                        k = 2)
# bidirectional model with BIC criterion
bidirectional_model_BIC = stepAIC(full_model, 
                        scope = list(upper = formula(full_model),
                                     lower = formula(null_model)),
                        direction = 'both',
                        k = log(n))
```

We see that with all three methods we converge onto the model 

$$fatalities \thicksim speed70 + alcohol + income + age$$

If we consider a polynomial term we seem to get substationally better: 

```{r}
quadratic_model = lm(fatalities ~ speed70 + alcohol + age + income + I(income^2), data = df.no_ss)
summary(quadratic_model)
```
$$fatalities \thicksim speed70 + alcohol + age + income + 
    I(income^2)$$

The final thing we probably want to consider is decreasing the categorical variables: 

1. take away speed70

```{r}
no_speed70_model = lm(fatalities ~ alcohol + age + income + I(income^2), data = df.no_ss)
summary(no_speed70_model)
```

2. take away alcohol

```{r}
no_alcohol_model = lm(fatalities ~ speed70 + age + income + I(income^2), data = df.no_ss)
summary(no_alcohol_model)
```

3. take away both

```{r}
continous_model = lm(fatalities ~ age + income + I(income^2), data = df.no_ss)
summary(continous_model)
```

seems like $$fatalities \thicksim speed70 + age + income + I(income^2)$$
could be the best (striking a good balance between everything)

## DO IT ALL AGAIN WITH THE POLYNOMIAL TERM
```{r}
#df_no_na = na.omit(df)
df_no_na = df %>% mutate(seatbelt = case_when(is.na(seatbelt) ~ 1, TRUE ~ seatbelt))
```

```{r}
# take away state and seatbelt as variables and create the full and null model
full_model = lm(fatalities ~ . + I(income^2), data = df_no_na)
null_model = lm(fatalities ~ 1 , data = df_no_na)
```


we perform a forward search with the aic and bic criterion
```{r}
# forward model with aic criterion
forward_model_aic = stepAIC(null_model, 
                        scope = list(upper = formula(full_model),
                                     lower = formula(null_model)),
                        direction = 'forward',
                        k = 2)
# forward model with bic criterion
forward_model_bic = stepAIC(null_model, 
                        scope = list(upper = formula(full_model),
                                     lower = formula(null_model)),
                        direction = 'forward',
                        k = log(n))
```
model for forward aic: 
$$fatalities \thicksim state + year + miles + I(income^2) + 
    speed70 + age + seatbelt + enforce$$

model for forward bic: 
$$fatalities \thicksim income + state + year + miles + speed70 + 
    age + seatbelt$$

```{r}
# backward model with aic criterion
backward_model_aic = stepAIC(full_model, 
                        scope = list(upper = formula(full_model),
                                     lower = formula(null_model)),
                        direction = 'backward',
                        k = 2)
# backward model with bic criterion
backward_model_bic = stepAIC(full_model, 
                        scope = list(upper = formula(full_model),
                                     lower = formula(null_model)),
                        direction = 'backward',
                        k = log(n))
```
Model for backward aic:
$$fatalities \thicksim state + year + miles + seatbelt + speed70 + 
    income + age + enforce$$
model for backward bic: 
$$fatalities \thicksim state + year + miles + seatbelt + speed70 + 
    income + age$$
```{r}
# bidirectional model with AIC criterion
bidirectional_model_aic = stepAIC(full_model, 
                        scope = list(upper = formula(full_model),
                                     lower = formula(null_model)),
                        direction = 'both',
                        k = 2)
# bidirectional model with BIC criterion
bidirectional_model_bic = stepAIC(full_model, 
                        scope = list(upper = formula(full_model),
                                     lower = formula(null_model)),
                        direction = 'both',
                        k = log(n))
```
Model for bidirectional aic: 
$$fatalities \thicksim state + year + miles + seatbelt + speed70 + 
    age + enforce + I(income^2)$$

Model for bidirectional bic: 
$$fatalities \thicksim state + year + miles + speed70 + age + 
    I(income^2)$$
## Model Building


### Variable Selection

* Interaction terms (from literature), Quadratic terms (())
* Forward, Backward, Bidirectional Search
* Innovation mark could be using AIC, BIC, Cp, Adjusted R squared
* Two models could be a larger subset model and a more parsimonious model with less variables with slightly lower metrics

### Inferences

* Confidence Interval & Hypothesis Testing
* **Question, do we need a hypothesis test for every single coefficient**

### Unusual Observations

* Leverage
* Outlier
* High Influence
* Must be repeated from each of models

### Checking Assumptions

* Check variance inflation factor below 10 for innovation mark
* Linearity Assumption
* Normality assumption (QQ plot + residual)

### Model Evaluation & Comparison

* **What is the prediction metric**
* Information Criteria
* k-fold Cross-Validation 

### Model Comparison


* As above





### References
- https://trid.trb.org/view/352138
- https://injuryprevention.bmj.com/content/6/2/109
- https://pubmed.ncbi.nlm.nih.gov/18242744/
- https://web.stanford.edu/~leinav/pubs/RESTAT2003.pdf

- https://www.medicine.mcgill.ca/epidemiology/joseph/courses/EPIB-621/interaction.pdf
- https://quantifyinghealth.com/why-and-when-to-include-interactions-in-a-regression-model/#:~:text=In%20a%20regression%20model%2C%20consider,been%20proven%20in%20previous%20studies.
- https://stats.stackexchange.com/questions/340009/how-to-know-which-interaction-terms-to-include-in-a-regression-model
- https://statisticsbyjim.com/regression/multicollinearity-in-regression-analysis/
- https://statisticsbyjim.com/regression/interaction-effects/